---
title: "lab3"
author: "wenhui li"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
dat <- readxl::read_xls('lab3.xls')
```
# 1
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
datclean <- dat %>% mutate(HighestSchoolCompleted = as.numeric(HighestSchoolCompleted),
                           AgeOfRespondent = as.numeric(AgeOfRespondent))
q1 <- datclean %>% group_by(MaritalStatus) %>% summarize(school = mean(HighestSchoolCompleted, na.rm=TRUE)) %>% arrange(desc(school))
ggplot(q1, aes(x= MaritalStatus,weight= school)) + geom_bar() + ylab('Average years of education')
```

From the barplot I would say that the average years of education have no relationship with the marital status.

# 2
```{r}
q2 <- filter(datclean, MaritalStatus == 'No answer' & HighestSchoolCompleted>15.5 | MaritalStatus == 'Divorced' & HighestSchoolCompleted >13.1 | MaritalStatus == 'Married' & HighestSchoolCompleted >13.0 | MaritalStatus == 'Never married' & HighestSchoolCompleted>13.4 | MaritalStatus == 'Separated' & HighestSchoolCompleted >11.9 | MaritalStatus == 'Widowed' & HighestSchoolCompleted >11.2)
str(q2)
```

# 3
```{r}
q3 <- filter(datclean, GeneralHappiness %in% c('Very happy','Pretty happy','Not too happy')) %>% mutate(HappyOrNot  = (GeneralHappiness %in% c('Very happy','Pretty happy')))
ggplot(q3, aes(x=MaritalStatus,fill = HappyOrNot)) + geom_bar(position ='fill')
```

From the barchart I would see that in all different kinds of marital status, nearly 75% of the respondents are happy. In my opinion, happiness has a very weak relationship with marital status.

# 4
```{r}
q4 <- filter(datclean, GeneralHappiness %in% c('Very happy','Pretty happy','Not too happy')) %>% mutate(HappyOrNot  = (GeneralHappiness %in% c('Very happy','Pretty happy')))
ggplot(q4, aes(x=PoliticalPartyAffiliation,fill = HappyOrNot)) + geom_bar()+coord_flip()
```

From the barplot I would say that respondents who have a political party are more happy than unhappy because there are more count of true than false in each bar.

# 5
## Does average age of the respondents affect the relationship with happiness?
```{r}
q5 <- filter(datclean, GeneralHappiness %in% c('Very happy','Pretty happy','Not too happy')) %>% mutate(HappyOrNot  = (GeneralHappiness %in% c('Very happy','Pretty happy')))
q5 <- q5 %>% group_by(HappyOrNot) %>% summarize(age = mean(AgeOfRespondent, na.rm=TRUE)) %>% arrange(desc(age))
table(q5)
```

From the table above, I would say that happiness has a weak relationship with average age. It is because the average age of happy people and unhappy people are all around 45-46. It's just a characteristic of respondents.